### 2024-04-30 paper.tex
use pdf, include references

\usepackage[utf8]{inputenc}
\usepackage[super,sort]{natbib}
%\bibliographystyle{simon}
\bibliographystyle{unsrtnat}
\setlength{\bibsep}{0.3ex plus 0.2ex}
%\setcitestyle{aysep={}} 
\usepackage[dvipsnames]{xcolor}
\usepackage[compact]{titlesec}
\usepackage[hypertexnames=false]{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={blue!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!50!black}
}
\urlstyle{same} % no tt font for URLs
\renewcommand{\topfraction}{0.9}
\renewcommand{\dbltopfraction}{0.9}
\renewcommand{\textfraction}{0.1}
\clubpenalty=1000
\widowpenalty=1000
\displaywidowpenalty=1000

@@

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=0.7\textwidth]{figures/Fig_residuals_AB.png}\\
	\end{center}
	\caption{\small \textbf{Improved quantification of DNA methylation in a given genomic interval.}\\
		\textbf{(A)} Depicted is a genomic interval (between vertical blue lines) along a chromosome, for which DNA methylation is to be quantified.
		Two cells cover differing parts of the interval with one read each.
		If one simply counts for each cell which fraction of the covered CpG sites are methylated, one obtains very different values for the two cells.
		\textbf{(B)} By averaging each CpG site's methylation over all cells and subsequent smoothing, the thick black ``average methylation curve'' is obtained.
		To quantify the methylation of cell 1 from (A) relative to this average over all cells, we propose to use the cell's residuals to the smoothed curve (lengths of the vertical red lines) and take their average, counting residuals of methylated CpGs positive and residuals of unmethylated CpGs negative.}
	\label{fig:smoothres}
\end{figure*}

\begin{figure}[p]
	\begin{center}
		\includegraphics[width=.7\columnwidth]{figures/Fig_sliding.png}
	\end{center}
	\caption{\small \textbf{Finding variably methylated regions.}\\
		The chromosomes are divided up into overlapping windows (first five shown at the bottom), and for each window, the cells' methylation values are calculated as described and as depicted in Fig.~\ref{fig:smoothres}B.
		Then, the variance of these values is calculated (each point represents one of the overlapping windows), a threshold (dashed line) is chosen such that a chosen quantile of windows have a variance exceeding the threshold.
		Windows with above-threshold variance are merged if they overlap, yielding the ``variably methylated regions'' (VMRs).}
	\label{fig:vmr}
\end{figure}

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=\textwidth]{figures/Fig_correlation.png}
	\end{center}
	\caption{\small \textbf{Correlation of DNA methylation and gene expression}.\\
		\textbf{(A)} Distribution of Pearson correlations between gene expression and promoter methylation (black) and gene expression and methylation of the nearest VMR (red): for most promoters, correlation is weak.
		Promoters are defined as intervals \textpm2~kb around the TSS.
		\textbf{(B)} Mean methylation near the gene \textit{Htra1}.
		Cells are assigned to 4 groups based on \textit{Htra1} expression (group 0: cells with no \textit{Htra1} expression, groups 1-3: cells that express \textit{Htra1}, divided into three equally large groups with group 3 having the highest expression).
		Data from \citet{kremer_scnmt}.}
	\label{fig:correlation}
\end{figure*}

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{figures/Fig_benchmark.png}
	\end{center}
	\caption{\small \textbf{Benchmark of our methods on single-cell multi-omic data of cells of the murine forebrain}.\\
		\textbf{(A)} Cell labels based on clustering of single-cell transcriptomes from \citet{kremer_scnmt}.
		\textbf{(B, C)} Exemplary UMAPs obtained when analyzing the data set with conventional methods based on genome tiling (B) or our proposed methods (C).
		\textbf{(D, E)} Mean degree of cell type separation (neighbor score, E, computed in 15-dimensional PC space) obtained when analyzing single-cell methylomes with different combinations of methods.
		Either VMRs, ENCODE regulatory elements, 100~kb genomic tiles or promoter regions (TSS\textpm2kb) were subjected to iterative PCA and UMAP.
		CpG methylation in these intervals was either quantified by averaging (dotted lines) or using the shrunken mean of the residuals as proposed in this work (solid lines).
		The full 1566-cell data set was sub-sampled to simulate smaller data sets (x axis).
		A higher neighbor score implies better separation of cell types inferred from single-cell transcriptomes of the same cells (A).
		\textbf{(F)} Proportion of VMRs (blue dots) that have at least 1~bp overlap with other genomic features.
		The black range illustrates the minimum and maximum overlap observed for 100 reshufflings of the VMRs to randomly chosen positions.
	}
	\label{fig:score}
\end{figure*}

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=.8\textwidth]{figures/Fig_DMRs.png}
	\end{center}
	\caption{\small \textbf{Detection of differentially methylated regions (DMRs).}\\
		\textbf{(A)} To identify DMRs, the chromosome is first divided into windows overlapping by a small and fixed size.
		For each window, the cell’s methylation values are calculated as described and as depicted in Fig.~\ref{fig:smoothres}B.
		Then the t statistic of these values is obtained as a measure of methylation difference between two groups of cells that were selected by the user (each point represents one of the windows).
		Upper and lower thresholds (dashed lines) are chosen such that a defined quantile of windows have a t statistic above or below the threshold.
		Windows exceeding either threshold are merged if they overlap, yielding DMRs that are lowly methylated in either group of cells.
		\textbf{(B)} DMRs detected between 58 oligodendrocytes (ODs) and 130 neural stem cells (NSCs) from \citet{kremer_scnmt} (top) and DMRs detected in the same data with randomly permuted cell labels (bottom, used to estimate FDR and determine adjusted p-values).
		\textbf{(C)} Enrichment of GO terms associated with DMRs lowly methylated in ODs (top) or NSCs (bottom).
		Depicted are the top 5 GO terms of the "Biological Process" (GO:BP) and "Cellular Component" (GO:CC) GO category, and their binomial test p-value (two-sided, not adjusted for multiple comparisons) and enrichment, as reported by GREAT \citep{mclean2010great}.
		\textbf{(D)} Mean methylation of NSCs and ODs at an exemplary DMR.
		Each point corresponds to a cell, black lines denote the median.
		\textbf{(E)} Detailed view of the DMR (yellow dashed rectangle) from C in pseudo-bulk samples consisting of NSCs or ODs.
		Vertical bars represent CpG sites.
	}
	\label{fig:dmr}
\end{figure*}

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=.8\textwidth]{figures/Fig_workflow.png}
	\end{center}
	\caption{\small \textbf{Overview of the functionalities implemented in the \textit{MethSCAn} package.}\\
		Common bisulfite sequencing mappers produce one methylation report file per cell.
		As it is inconvenient to work with thousands of text files, \texttt{methscan prepare} first stores their contents in a compressed format optimized for efficient access to genomic intervals.
		This command also produces cell-wise summary statistics that can be used to discard low-quality cells with \texttt{methscan filter}.
		The quality of each cell can furthermore be assessed by visualizing the average methylation around transcription start sites, which are known to show a characteristic dip in mammalian genomes, using \texttt{methscan profile}.
		This command can also visualize methylation at other genomic features of interest, such as enhancers or transcription factor binding sites.
		To obtain a methylation matrix analogous to scRNA-seq count matrices, the user must first decide in which genomic intervals methylation should be quantified.
		The user may either provide genomic regions of \emph{a priori} interest as a BED file, or they may choose to discover VMRs (variably methylated regions) in the data with \texttt{methscan scan}.
		The \texttt{methscan matrix} command then quantifies methylation at these genomic intervals in all cells.
		It provides two metrics:
		the simple percentage of methylated CpG sites and our proposed measure, the shrunken mean of the residuals, which is more robust to variation in sequencing coverage and stochastic differences in read position between cells.
		The resulting methylation matrix can then serve as input for established methods used for the exploration of single-cell data, such as dimensionality reduction and cell clustering.
		Finally, after annotation and exploration of the data set, the user may specify two groups of cells for DMR (differentially methylated region) detection with \texttt{methscan diff}.
		This command produces a BED file listing the genome coordinates, adjusted p-values, as well as several other metrics associated with DMRs.
		These DMRs may then be associated with nearby genes, or used for GO enrichment with tools such as GREAT \citep{mclean2010great} to enable functional interpretation.
	}
	\label{fig:workflow}
\end{figure*}

@@

{\phantomsection\addcontentsline{toc}{subsection}{References}
\small \balance \bibliography{scbs} \balance}

\onecolumn 
\newpage
\section*{Extended Data Figures}
\vspace{1.5cm}

\setcounter{figure}{0}
%\renewcommand{\thefigure}{ \arabic{figure}}
\renewcommand{\figurename}{Extended Data Figure}

\noindent%
\begin{minipage}{\linewidth}
	\begin{center}
	    \includegraphics[width=\textwidth]{figures/SFig_benchmark.png}
	\end{center}
	\captionof{figure}{\small \textbf{Benchmark of several combinations of analysis methods on four single-cell methylation data sets.}\\
	\textbf{(A-E)} Benchmarking the ability to separate groups of cells based on CpG methylation data, using different analysis approaches.
	Methylation matrices were obtained either by averaging CpG methylation in genomic intervals (dotted lines, hollow points) or using the shrunken mean of residuals (solid lines and points).
	Each analysis was performed for the full data sets and sub-samples of each.
	Four sets of genomic intervals (VMRs, ENCODE regulatory elements, 100~kb tiles or promoters) were quantified and separately subjected to four dimensionality reduction techniques (see Methods for details):
	iterative PCA as proposed by us, mean-imputed PCA on intervals with high sequencing coverage \citep{luo2017single}, mean-imputed PCA on all intervals, and MOFA+ \citep{argelaguet2020mofa} on all intervals.
	For each combination of methods, we quantified the ability to use distance in 15-dimensional PCA space to separate ground-truth cell labels reported by the authors (neighbor score).
	\textbf{(A)} Separation of neural cell types/states annotated based on scRNA-seq of the same cells \citep{kremer_scnmt}.
	\textbf{(B)} Separation of neuron sub-types, annotated by averaging CH-methylation (mCH) in 100~kb genomic tiles \citep{luo2017single}.
	\textbf{(C)} Separation of human colorectal cancer sampling regions (primary tumor, normal adjacent tissue, lymph node metastasis, liver metastasis before and after treatment, omental metastasis) \citep{bian2018single}.
	\textbf{(D-E)} Separation of three properties of mouse embryo cells:
	The developmental stage (E4.5, E5.5, E6.5, E7.5), the germ layer or extra-embryonic tissue of origin, or the cell lineage.
	Germ layer and lineage (E, only E7.5-cells) were annotated by \citet{argelaguet2019gastru} based on scRNA-seq of the same cells.
	}
	\label{fig:score_luo}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[super,sort]{natbib}
\bibliographystyle{unsrtnat}
\setlength{\bibsep}{0.3ex plus 0.2ex}
\usepackage[dvipsnames]{xcolor}
\usepackage[compact]{titlesec}
\usepackage[hypertexnames=false]{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={blue!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!50!black}
}
\urlstyle{same} % no tt font for URLs
\renewcommand{\topfraction}{0.9}
\renewcommand{\dbltopfraction}{0.9}
\renewcommand{\textfraction}{0.1}
\clubpenalty=1000
\widowpenalty=1000
\displaywidowpenalty=1000

@@

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=0.7\textwidth]{figures/Fig_residuals_AB.pdf}\\
	\end{center}
	\caption{\small \textbf{Improved quantification of DNA methylation in a given genomic interval.}\\
		\textbf{(A)} Depicted is a genomic interval (between vertical blue lines) along a chromosome, for which DNA methylation is to be quantified.
		Two cells cover differing parts of the interval with one read each.
		If one simply counts for each cell which fraction of the covered CpG sites are methylated, one obtains very different values for the two cells.
		\textbf{(B)} By averaging each CpG site's methylation over all cells and subsequent smoothing, the thick black ``average methylation curve'' is obtained.
		To quantify the methylation of cell 1 from (A) relative to this average over all cells, we propose to use the cell's residuals to the smoothed curve (lengths of the vertical red lines) and take their average, counting residuals of methylated CpGs positive and residuals of unmethylated CpGs negative.}
	\label{fig:smoothres}
\end{figure*}

\begin{figure}[p]
	\begin{center}
		\includegraphics[width=.7\columnwidth]{figures/Fig_sliding.pdf}
	\end{center}
	\caption{\small \textbf{Finding variably methylated regions.}\\
		The chromosomes are divided up into overlapping windows (first five shown at the bottom), and for each window, the cells' methylation values are calculated as described and as depicted in Fig.~\ref{fig:smoothres}B.
		Then, the variance of these values is calculated (each point represents one of the overlapping windows), a threshold (dashed line) is chosen such that a chosen quantile of windows have a variance exceeding the threshold.
		Windows with above-threshold variance are merged if they overlap, yielding the ``variably methylated regions'' (VMRs).}
	\label{fig:vmr}
\end{figure}

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=\textwidth]{figures/Fig_correlation.pdf}
	\end{center}
	\caption{\small \textbf{Correlation of DNA methylation and gene expression}.\\
		\textbf{(A)} Distribution of Pearson correlations between gene expression and promoter methylation (black) and gene expression and methylation of the nearest VMR (red): for most promoters, correlation is weak.
		Promoters are defined as intervals \textpm2~kb around the TSS.
		\textbf{(B)} Mean methylation near the gene \textit{Htra1}.
		Cells are assigned to 4 groups based on \textit{Htra1} expression (group 0: cells with no \textit{Htra1} expression, groups 1-3: cells that express \textit{Htra1}, divided into three equally large groups with group 3 having the highest expression).
		Data from \citet{kremer_scnmt}.}
	\label{fig:correlation}
\end{figure*}

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{figures/Fig_benchmark.pdf}
	\end{center}
	\caption{\small \textbf{Benchmark of our methods on single-cell multi-omic data of cells of the murine forebrain}.\\
		\textbf{(A)} Cell labels based on clustering of single-cell transcriptomes from \citet{kremer_scnmt}.
		\textbf{(B, C)} Exemplary UMAPs obtained when analyzing the data set with conventional methods based on genome tiling (B) or our proposed methods (C).
		\textbf{(D, E)} Mean degree of cell type separation (neighbor score, E, computed in 15-dimensional PC space) obtained when analyzing single-cell methylomes with different combinations of methods.
		Either VMRs, ENCODE regulatory elements, 100~kb genomic tiles or promoter regions (TSS\textpm2kb) were subjected to iterative PCA and UMAP.
		CpG methylation in these intervals was either quantified by averaging (dotted lines) or using the shrunken mean of the residuals as proposed in this work (solid lines).
		The full 1566-cell data set was sub-sampled to simulate smaller data sets (x axis).
		A higher neighbor score implies better separation of cell types inferred from single-cell transcriptomes of the same cells (A).
		\textbf{(F)} Proportion of VMRs (blue dots) that have at least 1~bp overlap with other genomic features.
		The black range illustrates the minimum and maximum overlap observed for 100 reshufflings of the VMRs to randomly chosen positions.
	}
	\label{fig:score}
\end{figure*}

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=.8\textwidth]{figures/Fig_DMRs.pdf}
	\end{center}
	\caption{\small \textbf{Detection of differentially methylated regions (DMRs).}\\
		\textbf{(A)} To identify DMRs, the chromosome is first divided into windows overlapping by a small and fixed size.
		For each window, the cell’s methylation values are calculated as described and as depicted in Fig.~\ref{fig:smoothres}B.
		Then the t statistic of these values is obtained as a measure of methylation difference between two groups of cells that were selected by the user (each point represents one of the windows).
		Upper and lower thresholds (dashed lines) are chosen such that a defined quantile of windows have a t statistic above or below the threshold.
		Windows exceeding either threshold are merged if they overlap, yielding DMRs that are lowly methylated in either group of cells.
		\textbf{(B)} DMRs detected between 58 oligodendrocytes (ODs) and 130 neural stem cells (NSCs) from \citet{kremer_scnmt} (top) and DMRs detected in the same data with randomly permuted cell labels (bottom, used to estimate FDR and determine adjusted p-values).
		\textbf{(C)} Enrichment of GO terms associated with DMRs lowly methylated in ODs (top) or NSCs (bottom).
		Depicted are the top 5 GO terms of the "Biological Process" (GO:BP) and "Cellular Component" (GO:CC) GO category, and their binomial test p-value (two-sided, not adjusted for multiple comparisons) and enrichment, as reported by GREAT \citep{mclean2010great}.
		\textbf{(D)} Mean methylation of NSCs and ODs at an exemplary DMR.
		Each point corresponds to a cell, black lines denote the median.
		\textbf{(E)} Detailed view of the DMR (yellow dashed rectangle) from C in pseudo-bulk samples consisting of NSCs or ODs.
		Vertical bars represent CpG sites.
	}
	\label{fig:dmr}
\end{figure*}

\begin{figure*}[p]
	\begin{center}
		\includegraphics[width=.8\textwidth]{figures/Fig_workflow.pdf}
	\end{center}
	\caption{\small \textbf{Overview of the functionalities implemented in the \textit{MethSCAn} package.}\\
		Common bisulfite sequencing mappers produce one methylation report file per cell.
		As it is inconvenient to work with thousands of text files, \texttt{methscan prepare} first stores their contents in a compressed format optimized for efficient access to genomic intervals.
		This command also produces cell-wise summary statistics that can be used to discard low-quality cells with \texttt{methscan filter}.
		The quality of each cell can furthermore be assessed by visualizing the average methylation around transcription start sites, which are known to show a characteristic dip in mammalian genomes, using \texttt{methscan profile}.
		This command can also visualize methylation at other genomic features of interest, such as enhancers or transcription factor binding sites.
		To obtain a methylation matrix analogous to scRNA-seq count matrices, the user must first decide in which genomic intervals methylation should be quantified.
		The user may either provide genomic regions of \emph{a priori} interest as a BED file, or they may choose to discover VMRs (variably methylated regions) in the data with \texttt{methscan scan}.
		The \texttt{methscan matrix} command then quantifies methylation at these genomic intervals in all cells.
		It provides two metrics:
		the simple percentage of methylated CpG sites and our proposed measure, the shrunken mean of the residuals, which is more robust to variation in sequencing coverage and stochastic differences in read position between cells.
		The resulting methylation matrix can then serve as input for established methods used for the exploration of single-cell data, such as dimensionality reduction and cell clustering.
		Finally, after annotation and exploration of the data set, the user may specify two groups of cells for DMR (differentially methylated region) detection with \texttt{methscan diff}.
		This command produces a BED file listing the genome coordinates, adjusted p-values, as well as several other metrics associated with DMRs.
		These DMRs may then be associated with nearby genes, or used for GO enrichment with tools such as GREAT \citep{mclean2010great} to enable functional interpretation.
	}
	\label{fig:workflow}
\end{figure*}

@@

\begin{thebibliography}{31}
	\providecommand{\natexlab}[1]{#1}
	\providecommand{\url}[1]{\texttt{#1}}
	\expandafter\ifx\csname urlstyle\endcsname\relax
	\providecommand{\doi}[1]{doi: #1}\else
	\providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

	\bibitem[Frommer et~al.(1992)Frommer, McDonald, Millar, Collis, Watt, Grigg,
	Molloy, and Paul]{Frommer_1992}
	M.~Frommer, L.~E. McDonald, D.~S. Millar, C.~M. Collis, F.~Watt, G.~W. Grigg,
	P.~L. Molloy, and C.~L. Paul.
	\newblock A genomic sequencing protocol that yields a positive display of
	5-methylcytosine residues in individual {DNA} strands.
	\newblock \emph{Proceedings of the National Academy of Sciences}, 89\penalty0
	(5):\penalty0 1827--1831, mar 1992.
	\newblock \doi{10.1073/pnas.89.5.1827}.
	\newblock URL \url{https://doi.org/10.1073/pnas.89.5.1827}.

	\bibitem[Smallwood et~al.(2014)Smallwood, Lee, Angermueller, Krueger, Saadeh,
	Peat, Andrews, Stegle, Reik, and Kelsey]{Smallwood_2014}
	S{\'{e}}bastien~A Smallwood, Heather~J Lee, Christof Angermueller, Felix
	Krueger, Heba Saadeh, Julian Peat, Simon~R Andrews, Oliver Stegle, Wolf Reik,
	and Gavin Kelsey.
	\newblock Single-cell genome-wide bisulfite sequencing for assessing epigenetic
	heterogeneity.
	\newblock \emph{Nature Methods}, 11\penalty0 (8):\penalty0 817--820, jul 2014.
	\newblock \doi{10.1038/nmeth.3035}.
	\newblock URL \url{https://doi.org/10.1038/nmeth.3035}.

	\bibitem[Hu et~al.(2016)Hu, Huang, An, Du, Hu, Xue, Zhu, Wang, Xue, and
	Fan]{scMTseq}
	Youjin Hu, Kevin Huang, Qin An, Guizhen Du, Ganlu Hu, Jinfeng Xue, Xianmin Zhu,
	Cun-Yu Wang, Zhigang Xue, and Guoping Fan.
	\newblock Simultaneous profiling of transcriptome and {DNA} methylome from a
	single cell.
	\newblock \emph{Genome Biology}, 17\penalty0 (1):\penalty0 1--11, 2016.
	\newblock \doi{10.1007/978-1-4939-9240-9_21}.

	\bibitem[Clark et~al.(2018)Clark, Argelaguet, Kapourani, Stubbs, Lee,
	Alda-Catalinas, Krueger, Sanguinetti, Kelsey, Marioni, Stegle, and
	Reik]{Clark2018}
	Stephen~J. Clark, Ricard Argelaguet, Chantriolnt-Andreas Kapourani, Thomas~M.
	Stubbs, Heather~J. Lee, Celia Alda-Catalinas, Felix Krueger, Guido
	Sanguinetti, Gavin Kelsey, John~C. Marioni, Oliver Stegle, and Wolf Reik.
	\newblock {scNMT}-seq enables joint profiling of chromatin accessibility {DNA}
	methylation and transcription in single cells.
	\newblock \emph{Nature Communications}, 9\penalty0 (1), February 2018.
	\newblock \doi{10.1038/s41467-018-03149-4}.
	\newblock URL \url{https://doi.org/10.1038/s41467-018-03149-4}.

	\bibitem[Hao et~al.(2023)Hao, Stuart, Kowalski, Choudhary, Hoffman, Hartman,
	Srivastava, Molla, Madad, Fernandez-Granda, and Satija]{seurat5}
	Yuhan Hao, Tim Stuart, Madeline~H Kowalski, Saket Choudhary, Paul Hoffman,
	Austin Hartman, Avi Srivastava, Gesmira Molla, Shaista Madad, Carlos
	Fernandez-Granda, and Rahul Satija.
	\newblock Dictionary learning for integrative, multimodal and scalable
	single-cell analysis.
	\newblock \emph{Nature Biotechnology}, 2023.
	\newblock \doi{10.1038/s41587-023-01767-y}.

	\bibitem[Wolf et~al.(2018)Wolf, Angerer, and Theis]{Wolf_2018}
	F.~Alexander Wolf, Philipp Angerer, and Fabian~J. Theis.
	\newblock {SCANPY}: large-scale single-cell gene expression data analysis.
	\newblock \emph{Genome Biology}, 19:\penalty0 15, 2018.
	\newblock \doi{10.1186/s13059-017-1382-0}.

	\bibitem[Luo et~al.(2017)Luo, Keown, Kurihara, Zhou, He, Li, Castanon, Lucero,
	Nery, Sandoval, et~al.]{luo2017single}
	Chongyuan Luo, Christopher~L Keown, Laurie Kurihara, Jingtian Zhou, Yupeng He,
	Junhao Li, Rosa Castanon, Jacinta Lucero, Joseph~R Nery, Justin~P Sandoval,
	et~al.
	\newblock Single-cell methylomes identify neuronal subtypes and regulatory
	elements in mammalian cortex.
	\newblock \emph{Science}, 357\penalty0 (6351):\penalty0 600--604, 2017.
	\newblock \doi{10.1126/science.aan3351}.

	\bibitem[Bird(1986)]{bird1986cpg}
	Adrian~P Bird.
	\newblock {CpG-rich} islands and the function of {DNA} methylation.
	\newblock \emph{Nature}, 321\penalty0 (6067):\penalty0 209--213, 1986.
	\newblock \doi{10.1038/321209a0}.

	\bibitem[Argelaguet et~al.(2019)Argelaguet, Clark, Mohammed, Stapel, Krueger,
	Kapourani, Imaz-Rosshandler, Lohoff, Xiang, Hanna,
	et~al.]{argelaguet2019gastru}
	Ricard Argelaguet, Stephen~J Clark, Hisham Mohammed, L~Carine Stapel, Christel
	Krueger, Chantriolnt-Andreas Kapourani, Ivan Imaz-Rosshandler, Tim Lohoff,
	Yunlong Xiang, Courtney~W Hanna, et~al.
	\newblock Multi-omics profiling of mouse gastrulation at single-cell
	resolution.
	\newblock \emph{Nature}, 576\penalty0 (7787):\penalty0 487--491, 2019.
	\newblock \doi{10.1038/s41586-019-1825-8}.

	\bibitem[Kremer et~al.(2022)Kremer, Cerrizuela, Al~Shukairi, Ellinger, Straub,
	Dehler, Korkmaz, Weichenhan, Plass, Anders, and
	Martin-Villalba]{kremer_scnmt}
	Lukas P.~M. Kremer, Santiago Cerrizuela, Mohammad~E Al~Shukairi, Tobias
	Ellinger, Jannes Straub, Sascha Dehler, Aylin Korkmaz, Dieter Weichenhan,
	Christoph Plass, Simon Anders, and Ana Martin-Villalba.
	\newblock Single-cell triple-omics uncovers {DNA} methylation as key feature of
	stemness in the healthy and ischemic adult brain.
	\newblock \emph{bioRxiv preprint}, 2022.
	\newblock \doi{10.1101/2022.07.13.499860}.

	\bibitem[Moore et~al.(2020)Moore, Purcaro, Pratt, Epstein, Shoresh, Adrian,
	Kawli, Davis, Dobin, et~al.]{encode2020expanded}
	Jill~E Moore, Michael~J Purcaro, Henry~E Pratt, Charles~B Epstein, Noam
	Shoresh, Jessika Adrian, Trupti Kawli, Carrie~A Davis, Alexander Dobin,
	et~al.
	\newblock Expanded encyclopaedias of {DNA} elements in the human and mouse
	genomes.
	\newblock \emph{Nature}, 583\penalty0 (7818):\penalty0 699--710, 2020.
	\newblock \doi{10.1038/s41586-020-2493-4}.

	\bibitem[Bian et~al.(2018)Bian, Hou, Zhou, Li, Yong, Wang, Wang, Yan, Hu, Guo,
	et~al.]{bian2018single}
	Shuhui Bian, Yu~Hou, Xin Zhou, Xianlong Li, Jun Yong, Yicheng Wang, Wendong
	Wang, Jia Yan, Boqiang Hu, Hongshan Guo, et~al.
	\newblock Single-cell multiomics sequencing and analyses of human colorectal
	cancer.
	\newblock \emph{Science}, 362\penalty0 (6418):\penalty0 1060--1063, 2018.
	\newblock \doi{10.1126/science.aao3791}.

	\bibitem[Argelaguet et~al.(2020)Argelaguet, Arnol, Bredikhin, Deloro, Velten,
	Marioni, and Stegle]{argelaguet2020mofa}
	Ricard Argelaguet, Damien Arnol, Danila Bredikhin, Yonatan Deloro, Britta
	Velten, John~C Marioni, and Oliver Stegle.
	\newblock {MOFA+}: a statistical framework for comprehensive integration of
	multi-modal single-cell data.
	\newblock \emph{Genome Biology}, 21\penalty0 (1):\penalty0 1--17, 2020.
	\newblock \doi{10.1186/s13059-020-02015-1}.

	\bibitem[Liu et~al.(2021)Liu, Zhou, Tian, Luo, Bartlett, Aldridge, Lucero,
	Osteen, Nery, Chen, et~al.]{liu2021dna}
	Hanqing Liu, Jingtian Zhou, Wei Tian, Chongyuan Luo, Anna Bartlett, Andrew
	Aldridge, Jacinta Lucero, Julia~K Osteen, Joseph~R Nery, Huaming Chen, et~al.
	\newblock {DNA} methylation atlas of the mouse brain at single-cell resolution.
	\newblock \emph{Nature}, 598\penalty0 (7879):\penalty0 120--128, 2021.
	\newblock \doi{10.1038/s41586-020-03182-8}.

	\bibitem[Hebestreit et~al.(2013)Hebestreit, Dugas, and Klein]{Hebestreit2013}
	Katja Hebestreit, Martin Dugas, and Hans-Ulrich Klein.
	\newblock Detection of significantly differentially methylated regions in
	targeted bisulfite sequencing data.
	\newblock \emph{Bioinformatics}, 29\penalty0 (13):\penalty0 1647--1653, May
	2013.
	\newblock \doi{10.1093/bioinformatics/btt263}.
	\newblock URL \url{https://doi.org/10.1093/bioinformatics/btt263}.

	\bibitem[Korthauer et~al.(2018)Korthauer, Chakraborty, Benjamini, and
	Irizarry]{dmrseq}
	Keegan Korthauer, Sutirtha Chakraborty, Yuval Benjamini, and Rafael~A Irizarry.
	\newblock Detection and accurate false discovery rate control of differentially
	methylated regions from whole genome bisulfite sequencing.
	\newblock \emph{Biostatistics}, 20\penalty0 (3):\penalty0 367--383, February
	2018.
	\newblock \doi{10.1093/biostatistics/kxy007}.
	\newblock URL \url{https://doi.org/10.1093/biostatistics/kxy007}.

	\bibitem[McLean et~al.(2010)McLean, Bristor, Hiller, Clarke, Schaar, Lowe,
	Wenger, and Bejerano]{mclean2010great}
	Cory~Y McLean, Dave Bristor, Michael Hiller, Shoa~L Clarke, Bruce~T Schaar,
	Craig~B Lowe, Aaron~M Wenger, and Gill Bejerano.
	\newblock {GREAT} improves functional interpretation of cis-regulatory regions.
	\newblock \emph{Nature Biotechnology}, 28\penalty0 (5):\penalty0 495--501,
	2010.
	\newblock \doi{10.1038/nbt.1630}.

	\bibitem[Boggs(2006)]{mbp}
	JM~Boggs.
	\newblock Myelin basic protein: a multifunctional protein.
	\newblock \emph{Cellular and Molecular Life Sciences CMLS}, 63:\penalty0
	1945--1961, 2006.
	\newblock \doi{10.1007/s00018-006-6094-7}.

	\bibitem[Krueger and Andrews(2011)]{bismark}
	Felix Krueger and Simon~R Andrews.
	\newblock Bismark: a flexible aligner and methylation caller for
	{Bisulfite-Seq} applications.
	\newblock \emph{Bioinformatics}, 27\penalty0 (11):\penalty0 1571--1572, 2011.
	\newblock \doi{10.1093/bioinformatics/btr167}.

	\bibitem[Schultz et~al.(2015)Schultz, He, Whitaker, Hariharan, Mukamel, Leung,
	Rajagopal, Nery, Urich, Chen, et~al.]{methylpy}
	Matthew~D Schultz, Yupeng He, John~W Whitaker, Manoj Hariharan, Eran~A Mukamel,
	Danny Leung, Nisha Rajagopal, Joseph~R Nery, Mark~A Urich, Huaming Chen,
	et~al.
	\newblock Human body epigenome maps reveal noncanonical {DNA} methylation
	variation.
	\newblock \emph{Nature}, 523\penalty0 (7559):\penalty0 212--216, 2015.
	\newblock \doi{10.1038/nature14465}.

	\bibitem[Zhou et~al.(2021)]{biscuit}
	Wanding Zhou et~al.
	\newblock {BISCUIT} -- understand sequencing data with bisulfite conversion.
	\newblock Software, \url{https://huishenlab.github.io/biscuit}, 2021.

	\bibitem[Kapourani and Sanguinetti(2019)]{kapourani2019melissa}
	Chantriolnt-Andreas Kapourani and Guido Sanguinetti.
	\newblock Melissa: Bayesian clustering and imputation of single-cell
	methylomes.
	\newblock \emph{Genome Biology}, 20\penalty0 (1):\penalty0 61, 2019.
	\newblock \doi{10.1186/s13059-019-1665-8}.

	\bibitem[Kapourani et~al.(2021)Kapourani, Argelaguet, Sanguinetti, and
	Vallejos]{kapourani2021scmet}
	Chantriolnt-Andreas Kapourani, Ricard Argelaguet, Guido Sanguinetti, and
	Catalina~A Vallejos.
	\newblock {scMET}: Bayesian modeling of {DNA} methylation heterogeneity at
	single-cell resolution.
	\newblock \emph{Genome Biology}, 22:\penalty0 1--21, 2021.
	\newblock \doi{10.1186/s13059-021-02329-8}.

	\bibitem[Danese et~al.(2021)Danese, Richter, Chaichoompu, Fischer, Theis, and
	Colomé-Tatché]{danese2021episcanpy}
	Anna Danese, Maria~L. Richter, Kridsadakorn Chaichoompu, David~S. Fischer,
	Fabian~J. Theis, and Maria Colomé-Tatché.
	\newblock {EpiScanpy}: integrated single-cell epigenomic analysis.
	\newblock \emph{Nature Communications}, 12:\penalty0 5228, 2021.
	\newblock \doi{10.1038/s41467-021-25131-3}.

	\bibitem[Welch et~al.(2019)Welch, Kozareva, Ferreira, Vanderburg, Martin, and
	Macosko]{welch2019single}
	Joshua~D Welch, Velina Kozareva, Ashley Ferreira, Charles Vanderburg, Carly
	Martin, and Evan~Z Macosko.
	\newblock Single-cell multi-omic integration compares and contrasts features of
	brain cell identity.
	\newblock \emph{Cell}, 177\penalty0 (7):\penalty0 1873--1887, 2019.
	\newblock \doi{10.1016/j.cell.2019.05.006}.

	\bibitem[Virtanen et~al.(2020)Virtanen, Gommers, Oliphant, Haberland, Reddy,
	et~al.]{SciPy}
	Pauli Virtanen, Ralf Gommers, Travis~E. Oliphant, Matt Haberland, Tyler Reddy,
	et~al.
	\newblock {SciPy 1.0: fundamental algorithms for scientific computing in
		Python}.
	\newblock \emph{Nature Methods}, 17:\penalty0 261--272, 2020.
	\newblock \doi{10.1038/s41592-019-0686-2}.

	\bibitem[Baglama and Reichel(2005)]{Baglama2005}
	James Baglama and Lothar Reichel.
	\newblock Augmented implicitly restarted {Lanczos} bidiagonalization methods.
	\newblock \emph{{SIAM} Journal on Scientific Computing}, 27\penalty0
	(1):\penalty0 19--42, January 2005.
	\newblock \doi{10.1137/04060593x}.
	\newblock URL \url{https://doi.org/10.1137/04060593x}.

	\bibitem[Josse and Husson(2012)]{josse2012}
	Julie Josse and Fran{\c c}ois Husson.
	\newblock {Handling missing values in exploratory multivariate data analysis
		methods}.
	\newblock \emph{{Journal de la Societe Fran{\c c}aise de Statistique}},
	153\penalty0 (2):\penalty0 79--99, 2012.
	\newblock URL \url{https://hal.science/hal-00811888}.
